Paper

### 2021-10-11 draft_paper.Rmd
addig plots to draft paper

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

@@

Overall volume of calls for service made in 2020 were below what we would expect based on past years' data. A higher proportion of calls were attended, and the response time was also shorter than the forecasts.

@@

- Drug related offences
- ASB - although COVID breach was coded as ASB so can we even say this??
- Police generated?

@@

- Residential burglary
- Theft from MV (and theft of)
- Theft other 
- Criminal damage (???)

@@

- Shoplifting
- Robbery

@@

Attend more stuff where less time at scene - eg night time economy calls takes people off scene for longer?

Officer workload changed (lowered) so more patrol and offering reassurance. Lower workload more responsive?

Budget for covid cars - also free up resources - these dealt with the ASB?

One explanation for why response time decreased?

@@

- highway disruption
- traffic collision
- Road Related Traffic Offence

Another explanation for why response time decreased?

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load libraries
library(lubridate)
library(tsibble)
library(tidyverse)

# Load calls (find own path to UoM server)
# calls <- read_csv("Z:\\n8_data_v2.csv.gz") %>%  janitor::clean_names()
calls <- read_csv("/Volumes/n8_covid/n8_data_v2.csv.gz") %>%  janitor::clean_names()

# Clean calls
# Find categories with fewer than a certain number of calls so we can exclude
# these, since these call types are not amenable to these types of incidents
minor_categories <- calls %>% 
  count(incident_type) %>% 
  filter(n < 1000) %>% 
  pull(incident_type)

# Categorise calls
calls <- mutate(
  calls,
  incident_type_new = case_when(
    incident_type %in% c("Distraction Burglary") ~
      "Burglary - Residential",
    incident_type %in% c(
      "Assistance to Other Agencies",
      "CRB Use Only",
      "CSI To Be Informed",
      "External System",
      "Found Stolen Vehicle",
      "Message to Pass/OOF Enquiries",
      "PNC Markers",
      "Police Vehicle Recovery",
      "Pre-Planned Events",
      "Stop Search",
      "Training"
    ) ~ "other - admin",
    incident_type %in% c(
      "Alarm - Activation", 
      "Alarm - No Response", 
      "Audible Only Alarm",
      "Insecure Premises",
      "Police Installed Alarm",
      "Unauthorised Encampment"
    ) ~ "alarms",
    incident_type %in% c(
      "Bail Breaches/Wanted Person",
      "Bail/Curfew/Wanted",
      "Prison Licence Recall",
      "Warrant Crown Court"
    ) ~ "warrant/bail issue",
    incident_type %in% c("Domestic Animal Concern", "Wildlife Matters") ~ 
      "animals",
    incident_type %in% c(
      "Firearms - Non Notifiable Crime",
      "Firearms Involved Crime"
    ) ~ "firearms",
    incident_type %in% c("RTC", "RTC - Damage Only") ~ "traffic collision",
    incident_type %in% minor_categories ~ "other - minor",
    TRUE ~ incident_type
  )
)

# Note important dates
# Source: https://www.instituteforgovernment.org.uk/sites/default/files/timeline-lockdown-web.pdf
dates <- tribble(
  ~date, ~event,
  "2020-01-31", "first UK COVID case",
  "2020-03-23", "first lockdown begins",
  "2020-06-15", "first lockdown ends",
  "2020-11-05", "second lockdown begins",
  "2020-12-02", "second lockdown ends"
) %>% 
  mutate(
    date = as_date(yearweek(ymd(date))), 
    row = row_number(),
    label = str_glue("{row}. {event}")
  )

# read in forecasts
# path to UoM server different for everyone
call_forecasts <- read_rds("/Volumes/n8_covid/call_forecasts.Rds")
all_rt_forecast <- read_rds("/Volumes/n8_covid/all_resp_time_forecast.Rds")
by_type_rt_forecast <- read_rds("/Volumes/n8_covid/by_type_rt_forecasts.Rds")
by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/by_grade_rt_forecasts.Rds")
all_calls_attended_forecast <- read_rds("/Volumes/n8_covid/all_attended_forecasts.Rds")
by_type_attended_forecast <- read_rds("/Volumes/n8_covid/by_type_attended_forecasts.Rds")

# Chart code
# specify the forecast (the df of the forecast), the variable we're modelling ("incident_type_new", "initial_grade_of_response"), and the value we're filtering on (e.g. "ASB" or "Grade 1").

forecast_chart <- function (forecasts, variable, value) {

  # get forecast type to string
  df_name <- deparse(substitute(forecasts))

  # conditionally label y axis
  y_axis_label <- case_when(df_name %in% c("by_type_rt_forecast", 
                                           "by_grade_rt_forecast") ~ "Median response time (in minutes)", 
                         df_name %in% c("call_forecasts") ~ "Weekly count of calls for service", 
                         df_name %in% c("by_type_attended_forecast", "by_grade_attended_forecast") ~ "Weekly % of calls attended")

  # conditionally find actual value column
  actual_col_name <- case_when(df_name %in% c("by_type_rt_forecast", 
                                           "by_grade_rt_forecast") ~ "actual_rt", 
                         df_name %in% c("call_forecasts") ~ "actual_calls", 
                         df_name %in% c("by_type_attended_forecast", "by_grade_attended_forecast") ~ "actual_attended")

  forecasts %>% 
    filter(eval(as.symbol(variable)) == value) %>% 
    # Occasionally the ARIMA model may produce forecast confidence intervals
    # that are less than zero, which will not show up on the plot because we
    # have set the y axis to start at zero (to make changes in vertical position
    # on the axis representative of changes in magnitude). To deal with this we
    # will manually set the lower CI of the forecasts to be zero if it is less
    # than zero in the original forecast.
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, eval(as.symbol(actual_col_name)))) +
    geom_point(aes(incident_week, eval(as.symbol(actual_col_name)), fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
                 limits = as.Date(c("2020-01-06", "2020-12-21")))+
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = value,
      subtitle = "",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = y_axis_label,
      fill = "Actual response time significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

}

```

@@

Overall volume of calls for service made in 2020 were below what we would expect based on past years' data.

```{r}

# Total calls fig

```

A higher proportion of calls were attended, and the response time was also shorter than the forecasts.

```{r}

# % attended fig and response time fig

```

@@

- Drug related offences

```{r drugsfig}

# Drugs fig
forecast_chart(call_forecasts, "incident_type_new", "Drugs")

```

- ASB - although COVID breach was coded as ASB so can we even say this??

```{r asbfig}

# ASB fig
forecast_chart(call_forecasts, "incident_type_new", "ASB")

```

- Police generated?

```{r policegenfig}

# Police generated fig
forecast_chart(call_forecasts, "incident_type_new", "Police Generated")

```

@@

- Residential burglary

```{r resburgfig}

# Residential burglary fig
forecast_chart(call_forecasts, "incident_type_new", "Burglary - Residential")

```

- Theft from MV (and theft of)

```{r theftmv}

# Theft of and theft from MV fig
forecast_chart(call_forecasts, "incident_type_new", "Theft From Motor Vehicle")
forecast_chart(call_forecasts, "incident_type_new", "Theft Of Motor Vehicle")

```

- Theft other

```{r}

# Theft other fig
forecast_chart(call_forecasts, "incident_type_new", "Theft Other")

```

- Criminal damage (???)

```{r crimdamfig}

# Crim dam fig
forecast_chart(call_forecasts, "incident_type_new", "Criminal Damage")

```

@@

- Shoplifting

```{r shopliftingfig}

# Shoplifting fig
forecast_chart(call_forecasts, "incident_type_new", "Shoplifting")

```

- Robbery

```{r robberyfig}

# Robbery fig
forecast_chart(call_forecasts, "incident_type_new", "Robbery")

```

### Domestic incidents

This is worth highlighting as there was much discussion about increase of domestic incidents during the lockdowns. When we look at all domestic incidents as a whole we do not really see this.

```{r domesticfig}

# domestic fig
forecast_chart(call_forecasts, "incident_type_new", "Domestic Incident")

```

However when we break down by grade of call

```{r}

# Domestic by grade fig

```

Police response by grade

```{r}

# Attended by grade fig and response time by grade fig

```

@@

Attend more stuff where less time at scene - eg night time economy calls takes people off scene for longer? *check time on scene*

With fewer calls for *list everything that had fewer than pre-covid* it is likely that officer workload changed (i.e. lowered). This meant quicker response times for calls. Figure X shows median response times by grade. Response times decreased for grade 1, grade 2, and grade 3 calls, the most drastic decrease in the latter category. This suggests that the freed up resources allowed faster responding to routine response calls. There was no difference in response time to those grade 4 calls (these require attendance or other resolution within 48 hours) and grade 5 calls are not attended.

```{r}

# Figs for grade 1, 2, 3, and 4 calls response times
forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 1")
forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 2")
forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 3")
forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 4")

```

Budget for covid cars - also free up resources - these dealt with the ASB? **Ask in interview about "covid cars" and whether these dealt with ASB**

@@

Another possible explanation for the decrease in response times of course is the quiet streets. We see that calls related to roads all dropped off in the first lock down, as fewer people were driving. Figures A, B, and C show how calls for highway disruption, traffic collision, and road related traffic offences were all less than would be expected based on previous years' data.

```{r roadsetcfigs}

# Figs for highway disruption, traffic collision, and road related traffic offences
forecast_chart(call_forecasts, "incident_type_new" ,"traffic collision")
forecast_chart(call_forecasts, "incident_type_new" ,"Highway Disruption")
forecast_chart(call_forecasts, "incident_type_new" ,"Road Related Traffic Offence")

```
